oREV item selection

### 2022-11-18 analysis/irt_analysis.Rmd
advance item selection

```{r}
coef_irt3 <- coef(irt3)

@@

```{r}
icc3 %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  #geom_textline(aes(label = item)) +
  #geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  #guides(col = F)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 3PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

@@

>>>>

```{r}
coef_irt3 <- coef(irt3) # robust = T (default = F)

@@

```{r}
icc3 %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  geom_textline(aes(label = item)) +
  geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  #geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  #guides(col = F)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 3PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

@@

# Manually excluding items

```{r}
irt_dat_sel_man <- irt_dat_sel%>%
  filter(!targetWord %in% c("badger", "jaguar","satellite"))

irt1_guess_sel_man <- brm(
  data = irt_dat_sel_man,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (1 | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  prior = prior_1pl_guess,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000
)%>%add_criterion(c("loo","waic"))

irt3_sel_man <- brm(
  data = irt_dat_sel_man,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(exp(logalpha) * eta),
    eta ~ 1 + (1 |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (1 |i| targetWord),
    nl = TRUE
  ),
  prior = prior_va_3pl_sel,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000
)%>%add_criterion(c("loo","waic"))

loo_compare(irt1_guess_sel_man, irt3_sel_man)%>%as_tibble(rownames = "model")

### 2022-11-18 analysis/item_selection.Rmd
advance item selection

irt1_guess <- readRDS("../saves/irt1_guess.rds")
```

@@

```{r}
rasch_fit <- irt_dat%>%
  left_join(aoa)%>%
  select(-order, -sex)%>%
  filter(targetWord %in% rasch_selected_items,
         #subjID %in% c("15X7mwV1", "39tVizn3", "479FefBH")
         )%>%
  add_epred_draws(irt1_guess, re_formula = ~(1 | targetWord) + (1 | subjID), ndraws = 5000)%>%
  mutate(zvi = (correct - .epred)/(.epred*(1-.epred))^0.5)%>%
  group_by(targetWord,aoa_german_comb, .draw)%>%
  summarise(outfit = sum(zvi^2)/length(unique(subjID)),
            infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))

rasch_fit_mode <- rasch_fit%>%
  pivot_longer(names_to = "fit_index", values_to = "value", cols = c(outfit, infit))%>%
  group_by(targetWord, fit_index)%>%
  summarise(mode = estimate_mode(value),
            lci = hdi_lower(value),
            uci = hdi_upper(value))

@@

@@

```{r}
coef_rasch <- coef(irt1)

@@

items <- eta_rasch%>%pull(item)
```

```{r}
easiness_rasch <- eta_rasch %>% pull(Estimate.Intercept)
infit <- rasch_fit_mode %>% filter(fit_index == "infit")%>%pull(mode)
outfit <- rasch_fit_mode %>% filter(fit_index == "outfit")%>%pull(mode)

@@

@@

@@

@@

ggplot(outfit_dist, aes(x = factor(size), y = dist_to_1))+
  geom_point(pch = 1)+
  theme_minimal()

```{r}
score_fn <- function(subset) {
	easinesses <- sort(easiness[subset])
	nn_dists <- rep(0, sum(subset)-1)
	for(i in 1:sum(subset)-1) {
		nn_dists[i] <- easinesses[i+1] - easinesses[i]
	}
	spacing <- -1*sd(nn_dists)/10
	infit_sample <- infit[subset]
	infit_dist <- unlist(lapply(infit_sample, func))
  mean_infit <- -4*mean(infit_dist)

@@

	# sel <- items[subset == TRUE]
	# 
	# sub_dat <- data%>%
	#   filter(targetWord %in% sel)%>%
	#   group_by(subjID)%>%
	#   summarise(mean_sub = mean(correct))
	# 
	# cor <- cor(sub_dat$mean_sub,full$mean_full)
	# 
	# return(spacing + mean_discrim/6 + cor/3)
	return(spacing + mean_infit + mean_outfit)
}

@@

simulated_annealing_rasch <- function(k, cooling_ratio=0.999, reset_thresh=1000, break_thresh=10000) {
  easiness <- eta_rasch %>% pull(Estimate.Intercept)
  infit <- rasch_fit_mode %>% filter(fit_index == "infit")%>%pull(mode)
  outfit <- rasch_fit_mode %>% filter(fit_index == "outfit")%>%pull(mode)
	N <- length(easiness)

@@

```{r, message=F, warning=F, comment=F}
sim_rasch <- simulated_annealing_rasch(22)

@@

```{r}
sub_rasch <- data%>%
  filter(targetWord %in% sel_items_rasch)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

@@

@@

@@

@@

for (i in 1:20) {

   sim <- simulated_annealing(22)

@@

frq_sel_items_rasch <- item_sel_rasch%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  arrange(-n)%>%
  slice(1:22)%>%
  pull(targetWord)

@@

@@

rasch_fit_sel_guess <- irt_dat%>%
  left_join(aoa)%>%
  select(-order, -sex)%>%
  filter(targetWord %in% rasch_selected_items,
         #subjID %in% c("15X7mwV1", "39tVizn3", "479FefBH")
         )%>%
  add_epred_draws(irt1_guess_sel, re_formula = ~(1 | targetWord) + (1 | subjID), ndraws = 5000)%>%
  mutate(zvi = (correct - .epred)/(.epred*(1-.epred))^0.5)%>%
  group_by(targetWord,aoa_german_comb, .draw)%>%
  summarise(outfit = sum(zvi^2)/length(unique(subjID)),
            infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))

@@

>>>>

irt1_guess <- readRDS("../saves/irt1_guess.rds")

@@

```{r}
rasch_fit <- irt_dat%>%
  left_join(aoa)%>%
  select(-order, -sex)%>%
  filter(#targetWord %in% rasch_selected_items,
         #subjID %in% c("15X7mwV1", "39tVizn3", "479FefBH")
         )%>%
  add_epred_draws(irt1_guess, re_formula = ~(1 | targetWord) + (1 | subjID), ndraws = 5000)%>%
  mutate(zvi = (correct - .epred)/(.epred*(1-.epred))^0.5)%>%
  group_by(targetWord,aoa_german_comb, .draw)%>%
  summarise(outfit = sum(zvi^2)/length(unique(subjID)),
            infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))

rasch_fit_mode <- rasch_fit%>%
  pivot_longer(names_to = "fit_index", values_to = "value", cols = c(outfit, infit))%>%
  group_by(targetWord, aoa_german_comb, fit_index)%>%
  summarise(mode = estimate_mode(value),
            lci = hdi_lower(value),
            uci = hdi_upper(value))

@@

rasch_fit_mode%>%
  left_join(maxi%>%pivot_longer(cols = c(infit, outfit), names_to = "fit_index", values_to = "maxi_mode")%>%select(targetWord,influence_pareto_k, fit_index, maxi_mode))%>%
  ggplot(aes(y = fct_reorder(targetWord, aoa_german_comb)))+
  geom_point(aes(x = mode),pch = 1, alpha = .75, col = "dodgerblue")+
  geom_point(aes(x = maxi_mode),pch = 1, alpha = .75, col = "firebrick")+
  geom_point(aes(x = influence_pareto_k),pch = 1, alpha = .75, col = "darkgreen")+
  facet_grid(~fit_index)+
  theme_bw()

@@

```{r}
coef_rasch <- coef(irt1_guess)

@@

items <- eta_rasch%>%pull(item)

@@

```{r}
easiness_rasch <- eta_rasch %>% pull(Estimate.eta_Intercept)
infit <- rasch_fit_mode %>% filter(fit_index == "infit")%>%pull(mode)
outfit <- rasch_fit_mode %>% filter(fit_index == "outfit")%>%pull(mode)
disc_2PL <- coef_irt3$targetWord[, , "logalpha_Intercept"] %>% as_tibble(rownames = "item")%>%pull(Estimate)
```

@@

```{r}
disc_dist <- tibble()

for (i in 1:48) {
x <- sample(disc_2PL, i)
  x_dist <- unlist(lapply(x, func))
  var_x <- -1*var(x_dist)*25
  row <- tibble(iter = j,
                size = i, 
                dist_to_1 = var_x)
  disc_dist <- bind_rows(disc_dist,row)
}
}

ggplot(disc_dist, aes(x = factor(size), y = dist_to_1))+
  geom_point(pch = 1)+
  theme_minimal()

@@

```{r}
pkd_dist <- tibble()

for (j in 1:20) {

for (i in 1:48) {
x <- sample(pkd, i)
  mean_x <- mean(x)
  row <- tibble(iter = j,
                size = i, 
                pkd = mean_x)
  pkd_dist <- bind_rows(pkd_dist,row)
}
}

ggplot(pkd_dist, aes(x = factor(size), y = pkd))+
  geom_point(pch = 1)+
  theme_minimal()

@@

ggplot(outfit_dist, aes(x = factor(size), y = dist_to_1))+
  geom_point(pch = 1)+
  theme_minimal()
```

```{r}
score_fn <- function(subset) {
	easinesses <- sort(easiness[subset])
	nn_dists <- rep(0, sum(subset)-1)
	for(i in 1:sum(subset)-1) {
		nn_dists[i] <- easinesses[i+1] - easinesses[i]
	}
	spacing <- -1*sd(nn_dists)/5

	var_disc_sample <- disc_2PL[subset]
	var_disc_2PL <- -1*var(var_disc_sample)*5

	infit_sample <- infit[subset]
	infit_dist <- unlist(lapply(infit_sample, func))
  mean_infit <- -4*mean(infit_dist)

@@

  #mean_pkd <- mean(pkd[subset])/2

	# sel <- items[subset == TRUE]
	# 
	# sub_dat <- data%>%
	#   filter(targetWord %in% sel)%>%
	#   group_by(subjID)%>%
	#   summarise(mean_sub = mean(correct))
	# 
	# cor <- cor(sub_dat$mean_sub,full$mean_full)
	# 
	# return(spacing + mean_discrim/6 + cor/3)
	return(spacing + mean_infit + mean_outfit+ var_disc_2PL )
}

@@

simulated_annealing_rasch <- function(k, cooling_ratio=0.999, reset_thresh=1000, break_thresh=10000) {
  easiness_rasch <- eta_rasch %>% pull(Estimate.eta_Intercept)
  infit <- rasch_fit_mode %>% filter(fit_index == "infit")%>%pull(mode)
  #infit <- maxi%>%pull(infit)
  outfit <- rasch_fit_mode %>% filter(fit_index == "outfit")%>%pull(mode)
  #outfit <- maxi%>%pull(outfit)
  #pkd <- maxi%>%arrange(targetWord)%>%pull(influence_pareto_k)
  N <- length(easiness)

@@

```{r, message=F, warning=F, comment=F}
sim_rasch <- simulated_annealing_rasch(20)

@@

```{r}
sub_rasch <- data%>%
  filter(targetWord %in% sel_items_rasch,
         #!targetWord %in% c("badger", "jaguar","satellite")
         )%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

@@

  for (j in 1:1) {

@@

for (j in 1:1) {

for (i in 5:6) {

@@

```{r}
  irt_comp_size <- tibble()

  for(j in 1:5){

    sub_dat <- irt_dat%>%filter(targetWord %in% sel)

    m1PL <- update(irt1_guess, newdata =sub_dat, chains = 6, cores = 6, threads = threading(8), backend = "cmdstanr")%>%add_criterion(c("loo"))
  	m3PL <- update(irt3, newdata =sub_dat, chains = 6, cores = 6, threads = threading(8), backend = "cmdstanr")%>%add_criterion(c("loo"))

  	comp <- loo_compare(m1PL, m3PL)%>%as_tibble(rownames = "model")%>%filter(model == "m1PL")%>%mutate(ratio = abs(elpd_diff)/(2*se_diff))

  	sub_dat_cor <- sub_dat%>%
  	  group_by(subjID)%>%
  	  summarise(mean_sub = mean(correct))

  	cor <- cor(sub_dat_cor$mean_sub,full$mean_full)

  	row <- tibble(size = i, 
           iter = j,
           elpd_diff = as.numeric(comp%>%pull(elpd_diff)),
           se_diff = as.numeric(comp%>%pull(se_diff)),
           ratio = as.numeric(comp%>%pull(ratio)),
           correlation = cor)

    irt_comp_size <- bind_rows(irt_comp_size, row)
  }

#saveRDS(irt_comp_size, "../saves/irt_comp_size.rds")

#irt_comp_size <- readRDS("../saves/irt_comp_size.rds")

irt_comp_size%>%
  pivot_longer(cols = c(model_comparison_ratio, correlation), names_to = "type",values_to = "value")%>% 
  ggplot(aes(x = factor(size), y = value))+
  geom_point(pch = 1)+
  geom_line()+
  facet_grid(type~. , scales = "free_y")+
  theme_bw()+
  labs(x = "No. of items", y = "Comparison value")
```

@@

for (i in 1:10) {

   sim <- simulated_annealing(24)

@@

frq_sel_items_rasch <- item_sel_rasch%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  arrange(-n)%>%
  slice(1:24)%>%
  pull(targetWord)

@@

### ICC

```{r}
icc1_guess <- posterior_samples(irt1_guess)%>% 
  select(b_eta_Intercept, starts_with("r_targetWord"))%>%
  mutate(iter = 1:n()) %>% 
  pivot_longer(starts_with("r_targetWord"), names_to = "item", values_to = "xi") %>%
  mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
  expand(nesting(iter, b_eta_Intercept, item, xi),
         theta = seq(from = -6, to = 6, length.out = 100)) %>% 
  mutate(p = 0.25 + 0.75*inv_logit_scaled((b_eta_Intercept + theta + xi))) %>%  
  group_by(theta, item) %>% 
  summarise(p = mean(p))%>%
  left_join(aoa%>%rename(item = targetWord))
```

```{r}
icc1_guess %>% 
  filter(item %in% sel_items_rasch)%>%
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  #facet_wrap(~group)+
  #guides(col = F)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 1PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

@@

rasch_fit_guess <- irt_dat%>%
  left_join(aoa)%>%
  select(-order, -sex)%>%
  filter(targetWord %in% rasch_selected_items,
         #subjID %in% c("15X7mwV1", "39tVizn3", "479FefBH")
         )%>%
  add_epred_draws(irt1_guess, re_formula = ~(1 | targetWord) + (1 | subjID), ndraws = 5000)%>%
  mutate(zvi = (correct - .epred)/(.epred*(1-.epred))^0.5)%>%
  group_by(targetWord,aoa_german_comb, .draw)%>%
  summarise(outfit = sum(zvi^2)/length(unique(subjID)),
            infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))

@@

```{r}
rasch_fit%>%
  filter(targetWord %in% rasch_selected_items)%>%
  pivot_longer(names_to = "fit_index", values_to = "value", cols = c(outfit, infit))%>%
  ggplot(. , aes(y = fct_reorder(targetWord, aoa_german_comb), x = value, fill = fit_index, col = fit_index))+
  geom_vline(xintercept = c(0.5, 1.5), lty = 2, alpha = .5)+
  geom_vline(xintercept = c(0.7, 1.3), lty = 3, alpha = .5)+
  geom_vline(xintercept = 1, lty = 1, alpha = .5, col = "darkgreen")+
  stat_halfeye(alpha = .75, .width = c(0.66, 0.95), position = position_dodge(width = 0.5))+
  scale_fill_colorblind()+
  scale_color_colorblind()+
  labs(x = "Index value", y = "Target word (sorted by rated age of acquisition)")+
  #facet_grid(~fit_index)+
  xlim(0,3)+
  theme_bw()
```

```{r}
rasch_fit_sel_guess <- irt_dat%>%
  left_join(aoa)%>%
  select(-order, -sex)%>%
  filter(targetWord %in% rasch_selected_items,
         #subjID %in% c("15X7mwV1", "39tVizn3", "479FefBH")
         )%>%
  add_epred_draws(irt1_guess_sel, re_formula = ~(1 | targetWord) + (1 | subjID), ndraws = 5000)%>%
  mutate(zvi = (correct - .epred)/(.epred*(1-.epred))^0.5)%>%
  group_by(targetWord,aoa_german_comb, .draw)%>%
  summarise(outfit = sum(zvi^2)/length(unique(subjID)),
            infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))
```
